---
title: "NEV Fault Dataset Analysis"
format: html
jupyter: false
execute:
  echo: false
  warning: false
  message: false
css: styles.css # <--- ADD THIS LINE
---

```{r}
# ✅ Libraries
library(tidyverse)
library(plotly)
library(scales)
library(viridis) # viridis for color-safe blue-green

# ✅ Good theme, defined ONCE
# This theme is based on the one you previously confirmed works for your axis labels.
better_theme <- theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 16), # Explicit font size for title
    panel.grid.minor = element_blank()
  )

# ✅ Load data
# Ensure this path is correct for your system!
nev_raw <- read_csv("C:/Users/anshb/Downloads/NEV_fault_dataset new (1).csv",
                     col_types = cols(.default = "c"))

numeric_cols <- c("Voltage (V)", "Current (A)", "Motor Speed (RPM)",
                  "Temperature (°C)", "Vibration (g)",
                  "Ambient Temp (°C)", "Humidity (%)")

nev <- nev_raw %>%
  mutate(across(all_of(numeric_cols), ~ as.numeric(str_remove_all(., "\\."))))
```

```{r}
# --- Start of unnamed-chunk-2 (Corrected for color, less points, and scale labels) ---
# Compute means:
nev_summary <- nev %>%
  group_by(`Fault Label`) %>%
  summarise(
    avg_vibration = mean(`Vibration (g)`, na.rm = TRUE),
    avg_temp = mean(`Temperature (°C)`, na.rm = TRUE),
    avg_current = mean(`Current (A)`, na.rm = TRUE)
  )

# Prepare data for plotting: Sample a fraction of the data to reduce points
set.seed(123) # for reproducibility - good to keep for consistent sampling
nev_sampled <- nev %>%
  sample_frac(0.1) # Sample 10% of the data for fewer points

# Updated ggplot for p_better: Current vs Temperature
p_better <- ggplot(nev_sampled, aes(x = `Current (A)`, y = `Temperature (°C)`, color = factor(`Fault Label`))) +
  geom_point(alpha = 0.7, size = 1.5) + # Increased point size slightly for better visibility with fewer points
  scale_x_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_color_viridis_d(option = "D") + # Set color to "D" option for consistency with other plots
  labs(title = "Current vs. Temperature by Fault Label (Sampled Data)",
       x = "Current (A)", y = "Temperature (°C)", color = "Fault Label") +
  better_theme # Apply the consistent theme

plotly::ggplotly(p_better)
```

This plot maps current against temperature, categorized by fault type. It highlights specific current-temperature combinations where certain faults consistently appear, providing insight into their operational footprint.


```{r}
# --- Start of unnamed-chunk-3 (Motor Speed Boxplot - no changes requested, but verified clean) ---
p2_boxplot <- ggplot(nev, aes(x = factor(`Fault Label`),
                              y = `Motor Speed (RPM)`,
                              fill = factor(`Fault Label`))) +
  geom_boxplot(alpha = 0.8, outlier.color = "red", outlier.size = 1) +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = "Motor Speed Distribution by Fault Label",
    x = "Fault Label",
    y = "Motor Speed (RPM)"
  ) +
  better_theme +
  theme(legend.position = "none")

ggplotly(p2_boxplot)
```

This plot presents the range and central tendency of motor speed for each fault type. It clarifies how faults influence motor operation, revealing if a fault is associated with consistent, higher, lower, or more variable speeds.


```{r}
# --- Start of unnamed-chunk-4 (Vibration Histogram - no changes requested, but verified clean) ---
p3 <- ggplot(nev, aes(x = `Vibration (g)`, fill = factor(`Fault Label`))) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30, color = "black") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(labels = label_number(scale_cut = cut_short_scale(), accuracy = 0.01)) +
  labs(
    title = "Distribution of Vibration by Fault Label",
    x = "Vibration (g)",
    fill = "Fault Label"
  ) +
  better_theme

ggplotly(p3) %>% layout(legend = list(title = list(text = 'Fault Label')))
```

This histogram depicts the prevalent vibration levels for each fault type. It establishes which vibration intensities correlate with specific faults, aiding in understanding fault severity and unique vibrational characteristics.

```{r}
# --- Start of unnamed-chunk-5 (Ambient Temp Violin - no changes requested, but verified clean) ---
p4_violin <- ggplot(nev, aes(x = factor(`Fault Label`),
                             y = `Ambient Temp (°C)`,
                             fill = factor(`Fault Label`))) +
  geom_violin(alpha = 0.8, trim = TRUE) +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = "Ambient Temperature Distribution by Fault Label",
    x = "Fault Label",
    y = "Ambient Temperature (°C)"
  ) +
  better_theme +
  theme(legend.position = "none")

ggplotly(p4_violin)
```



This plot shows ambient temperature distributions across different fault types. It identifies if particular faults are more likely under certain environmental temperatures, indicating the influence of external conditions on fault manifestation.



**Final Summary**


*These interactive plots offer a comprehensive view of the NEV fault dataset, revealing critical relationships between operational parameters and various fault conditions. The visualizations highlight distinct patterns in current, temperature, motor speed, vibration, and ambient temperature as they correspond to different fault labels. This analysis provides valuable insights for developing more effective predictive maintenance strategies, enabling earlier fault detection, and improving overall electric vehicle reliability.*
