---
title: "NEV Fault Dataset Analysis"
format: html
jupyter: false
execute:
  echo: false
  warning: false
  message: false
css: styles.css # <--- ADD THIS LINE
---

```{r}
# ✅ Libraries
# -----------------------------------------------------------
# Why: Load libraries, set custom theme, and read data.
# If the path breaks, update it for your local machine.
# -----------------------------------------------------------
library(tidyverse)
library(plotly)
library(scales)
library(viridis) # viridis for color-safe blue-green

# ✅ Good theme, defined ONCE
# This theme is based on the one you previously confirmed works for your axis labels.

# Consistent custom ggplot2 theme
better_theme <- theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 16), # Explicit font size for title
    panel.grid.minor = element_blank()
  )

library(readr) # make sure you have this for parse_number

# Better: parse numbers safely, keeping decimals, stripping units
nev <- nev_raw %>%
  mutate(across(all_of(numeric_cols), ~ parse_number(.)))
```

This project examines faults in New Energy Vehicles (NEVs) using real-world data and interactive visuals. By analyzing patterns, frequencies, and correlations, it aims to uncover insights to improve maintenance and support better NEV design. The visualizations make complex fault data easier to explore and understand, highlighting common issues in electric and hybrid vehicles. This resource is intended for researchers, engineers, and anyone interested in the technical side of cleaner transportation.

```{r}
# --- Start of unnamed-chunk-2 (Corrected for color, less points, and scale labels) ---
# -----------------------------------------------------------
# Why: Explore how current and temperature relate by fault type.
# Sampling reduces clutter and keeps plot readable.
# -----------------------------------------------------------

# Compute means by fault label
# Compute means:
nev_summary <- nev %>%
  group_by(`Fault Label`) %>%
  summarise(
    avg_vibration = mean(`Vibration (g)`, na.rm = TRUE),
    avg_temp = mean(`Temperature (°C)`, na.rm = TRUE),
    avg_current = mean(`Current (A)`, na.rm = TRUE)
  )

# Prepare data for plotting: Sample a fraction of the data to reduce points
# Sample 10% for clarity
set.seed(123) # for reproducibility - good to keep for consistent sampling
nev_sampled <- nev %>%
  sample_frac(0.1) # Sample 10% of the data for fewer points
# Plot: Current vs. Temperature by Fault
# Updated ggplot for p_better: Current vs Temperature
p_better <- ggplot(nev_sampled, aes(x = `Current (A)`, y = `Temperature (°C)`, color = factor(`Fault Label`))) +
  geom_point(alpha = 0.7, size = 1.5) + # Increased point size slightly for better visibility with fewer points
  scale_x_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_color_viridis_d(option = "D") + # Set color to "D" option for consistency with other plots
  labs(title = "Current vs. Temperature by Fault Label (Sampled Data)",
       x = "Current (A)", y = "Temperature (°C)", color = "Fault Label", caption = "Data Source: New Energy Vehicles") +
  better_theme # Apply the consistent theme

plotly::ggplotly(p_better)
```

This plot maps current against temperature, categorized by fault type. It highlights specific current-temperature combinations where certain faults consistently appear, providing insight into their operational footprint.


```{r}
# --- Start of unnamed-chunk-3 (Motor Speed Boxplot - no changes requested, but verified clean) ---
# -----------------------------------------------------------
# Why: Show motor speed spread for each fault type.
# Boxplots clarify variation & outliers.
# -----------------------------------------------------------
p2_boxplot <- ggplot(nev, aes(x = factor(`Fault Label`),
                              y = `Motor Speed (RPM)`,
                              fill = factor(`Fault Label`))) +
  geom_boxplot(alpha = 0.8, outlier.color = "red", outlier.size = 1) +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = "Motor Speed Distribution by Fault Label",
    x = "Fault Label",
    y = "Motor Speed (RPM)", caption = "Data Source: New Energy Vehicles"
  ) +
  better_theme +
  theme(legend.position = "none")

ggplotly(p2_boxplot)
```

This plot presents the range and central tendency of motor speed for each fault type. It clarifies how faults influence motor operation, revealing if a fault is associated with consistent, higher, lower, or more variable speeds.


```{r}
# --- Start of unnamed-chunk-4 (Vibration Histogram - no changes requested, but verified clean) ---
# -----------------------------------------------------------
# Why: Show vibration intensity distribution for each fault.
# Helps link vibration levels to specific fault types.
# -----------------------------------------------------------

p3 <- ggplot(nev, aes(x = `Vibration (g)`, fill = factor(`Fault Label`))) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30, color = "black") +
  scale_fill_viridis_d(option = "D") +
  scale_x_continuous(labels = label_number(scale_cut = cut_short_scale(), accuracy = 0.01)) +
  labs(
    title = "Distribution of Vibration by Fault Label",
    x = "Vibration (g)",
    fill = "Fault Label", caption = "Data Source: New Energy Vehicles"
  ) +
  better_theme

ggplotly(p3) %>% layout(legend = list(title = list(text = 'Fault Label')))
```

This histogram depicts the prevalent vibration levels for each fault type. It establishes which vibration intensities correlate with specific faults, aiding in understanding fault severity and unique vibrational characteristics.

```{r}
# --- Start of unnamed-chunk-5 (Ambient Temp Violin - no changes requested, but verified clean) ---
# -----------------------------------------------------------
# Why: Show ambient temperature variation by fault label.
# Violin plot highlights environmental impact on faults.
# -----------------------------------------------------------
p4_violin <- ggplot(nev, aes(x = factor(`Fault Label`),
                             y = `Ambient Temp (°C)`,
                             fill = factor(`Fault Label`))) +
  geom_violin(alpha = 0.8, trim = TRUE) +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = "Ambient Temperature Distribution by Fault Label",
    x = "Fault Label",
    y = "Ambient Temperature (°C)", caption = "Data Source: New Energy Vehicles"
  ) +
  better_theme +
  theme(legend.position = "none")

ggplotly(p4_violin)
```



This plot shows ambient temperature distributions across different fault types. It identifies if particular faults are more likely under certain environmental temperatures, indicating the influence of external conditions on fault manifestation.



**Final Summary**


*These interactive plots show key relationships in the NEV fault dataset, highlighting how factors like current, temperature, motor speed, vibration, and ambient temperature relate to different faults. This helps reveal patterns that can support better predictive maintenance, earlier fault detection, and improved electric vehicle reliability.*
